EldersAIR initial wood moisture content data work

### 2021-03-15 merge_data_files.Rmd
data cleaning updates

---
title: 'EldersAIR: data merging for analysis dataset'
author: "Ethan Walker"
date: "Started 11 March 2021, Updated 11 March 2021"
output:
  powerpoint_presentation: default
---

@@

elders_pm_at_home_visit <- read_rds(paste0(file_path, "Output/pm_at_home_visit.rds")) 
```

@@

Merge PM and health data by sampling visit
```{r}
health_pm_visit <- elders_medvisit_clean %>% 
  full_join(elders_pm_at_home_visit,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_extra_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_hx_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment"))

#summary(health_pm_visit)

write_rds(health_pm_visit, paste0(file_path, "Output/elders_health_pm_visit_dataset.rds"))
```

>>>>

---
title: 'EldersAIR: data merging for analysis dataset'
author: "Ethan Walker"
date: "Started 11 March 2021, Updated 15 March 2021"
output:
  powerpoint_presentation: default
---

@@

elders_pm_at_home_visit <- read_rds(paste0(file_path, "Output/pm_at_home_visit.rds"))

elders_stove_grades_clean <- read_rds(paste0(file_path, "Output/elders_stove_grades_clean.rds"))

elders_stove_usage_clean <- read_rds(paste0(file_path, "Output/elders_stove_usage_clean.rds"))

elders_kw_report_clean <- read_rds(paste0(file_path, "Output/elders_kw_report_clean.rds"))

elders_filter_type <- read_rds(paste0(file_path, "Output/elders_filter_type.rds"))

elders_temp_rh_clean <- read_rds(paste0(file_path, "Output/elders_temp_rh_clean.rds"))

elders_home_char_clean <- read_rds(paste0(file_path, "Output/elders_home_char_clean.rds"))

elders_home_activity_clean <- read_rds(paste0(file_path, "Output/elders_home_activity_clean.rds"))

elders_moisture_clean <- read_rds(paste0(file_path, "Output/elders_moisture_clean.rds")) 
```

@@

Prep demographics data for merging
```{r}
demographics <- elders_demographics %>% 
  arrange(area, adult_id_char, winter_id) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  rename("1" = sampling_date,
         "2" = sampling_date2) %>% 
  pivot_longer(cols = c("1", "2"), names_to = "sampling_visit",
               values_to = "sampling_date") %>% 
  mutate(sampling_visit = as.factor(sampling_visit)) %>% 
  select(area, home, home_id_num, home_winter_id, adult_id_char, adult_id_num,
         winter_id, sampling_visit, sampling_date, treatment, intervention_date,
         intervention_time, hispanic:residents_under18, -age) %>% 
  filter(!is.na(sampling_date)) %>% 
  arrange(area, adult_id_char, winter_id, sampling_visit)
```

Merge PM, health, stove data 
One observation per sampling visit (up to 4 obs per adult/home)
```{r}
health_pm_visit <- elders_medvisit_clean %>% 
  full_join(elders_pm_at_home_visit,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(demographics,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_extra_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_health_hx_clean,
            by = c("area", "home", "home_id_num", "adult_id_char", "adult_id_num",
                   "home_winter_id", "winter_id", "treatment")) %>% 
  left_join(elders_stove_grades_clean,
            by = c("area", "home", "home_id_num", "treatment")) %>% 
  left_join(elders_stove_usage_clean,
            by = c("area", "home", "home_id_num", 
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_kw_report_clean,
            by = c("area", "home_winter_id")) %>% 
  left_join(elders_filter_type,
            by = c("area", "home_winter_id")) %>% 
  left_join(elders_temp_rh_clean,
            by = c("area", "home_winter_id", "sampling_visit")) %>% 
  left_join(elders_home_activity_clean,
            by = c("area", "home", "home_id_num", 
                   "home_winter_id", "winter_id", "treatment", "sampling_visit")) %>% 
  left_join(elders_home_char_clean,
            by = c("area", "home", "home_id_num", 
                   "home_winter_id", "winter_id", "treatment")) %>% 
  left_join(elders_moisture_clean,
            by = c("area", "home_winter_id", "sampling_visit"))

# summary(health_pm_visit)
```

Save final dataset
```{r}
write_rds(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.rds"))

write_csv(health_pm_visit, paste0(file_path, "Output/elders_cleaned_dataset_by_visit.csv"), na = " ")

# save variable names to use in data dictionary
save_var_names <- data.frame(t(health_pm_visit)) %>% 
  rownames_to_column() %>% 
  select(rowname)

write_csv(save_var_names, paste0(file_path, "Output/health_pm_visit_var_names.csv"))
```

### 2021-03-15 wood_moisture_cleaning.Rmd
data cleaning updates

---
title: "EldersAIR initial wood moisture content data work"
author: "Ethan Walker"
date: "Started 25 Feb 2020, Updated 25 Feb 2020"
output: pdf_document
---

@@

# Load, format, and save wood moisture data
```{r}
## Box location: Update Box location once final data loaded
np_moisture_initial <- read_xlsx("Input/NP/np_moisture_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NP") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  arrange(home_winter_id, moisture_date) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>%
  filter(!is.na(moisture_ave)) %>% 
  arrange(home_winter_id, moisture_date)

## Box location: Update Box location once final data loaded
nn_moisture_initial <- read_xlsx("Input/NN/nn_moisture_20200225.xlsx") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  arrange(home_winter_id, moisture_date) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>%
  filter(!is.na(moisture_ave)) %>% 
  arrange(home_winter_id, moisture_date)

# Combine locations and save
elders_moisture <- rbind(np_moisture_initial, nn_moisture_initial)

#write_rds(elders_moisture, "Output/elders_moisture.rds")
```

# Clean up all moisture data
```{r}
elders_moisture <- read_rds("Output/elders_moisture.rds")

elders_ids_linked <- read_rds("Output/elders_ids_linked.rds")

elders_moisture_clean <- elders_moisture %>% 
  left_join(elders_ids_linked, by = c("area", "home_winter_id")) %>% 
  filter(!is.na(home_winter_id)) %>% 
  # lower range on instrument = 5%
  # therefore, changing values < 5 to 5
  mutate(moisture_1 = if_else(moisture_1 < 5, 5, moisture_1),
         moisture_2 = if_else(moisture_2 < 5, 5, moisture_2),
         moisture_3 = if_else(moisture_3 < 5, 5, moisture_3),
         moisture_ave = if_else(moisture_ave < 5, 5, moisture_ave))

# write_rds(elders_moisture_clean, "Output/elders_moisture_clean.rds")

elders_moisture_clean <- read_rds("Output/elders_moisture_clean.rds")
```

>>>>

---
title: "EldersAIR initial wood moisture content data work"
author: "Ethan Walker"
date: "Started 25 Feb 2020, Updated 15 March 2021"
output: pdf_document
---

@@

```{r}
file_path <- c("C:/Users/ethan.walker/Box/Ethan Walker UM/R/EldersAIR/")

elders_demographics <- read_rds(paste0(file_path, "Output/elders_demographics.rds")) %>% 
  arrange(area, adult_id_char, winter_id) %>% 
  group_by(area, adult_id_char, winter_id) %>% 
  rename("1" = sampling_date,
         "2" = sampling_date2) %>% 
  pivot_longer(cols = c("1", "2"), names_to = "sampling_visit",
               values_to = "sampling_date") %>% 
  mutate(sampling_visit = as.factor(sampling_visit)) %>% 
  ungroup() %>% 
  select(area, home_winter_id, sampling_visit, sampling_date, ) %>% 
  distinct(area, home_winter_id, sampling_date, .keep_all = T) %>% 
  filter(!is.na(sampling_date)) 
```

# Load, format, and save wood moisture data
```{r}
## Box location: Update Box location once final data loaded
npt_moisture_initial <- read_xlsx(paste0(file_path, "Input/NP/moisture.xlsx"),
                                  na = "NULL") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NPT") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  arrange(home_winter_id, moisture_date) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>%
  filter(!is.na(moisture_ave)) %>% 
  arrange(home_winter_id, moisture_date)

## Box location: Update Box location once final data loaded
nn_moisture_initial <- read_xlsx(paste0(file_path, "Input/NN/moisture.xlsx"),
                                  na = "NULL") %>% 
  # rename variables
  mutate(home_winter_id = as.factor(HomeWinterID),
         moisture_date = ymd(CompletedDate),
         moisture_1 = as.numeric(Moisture1),
         moisture_2 = as.numeric(Moisture2),
         moisture_3 = as.numeric(Moisture3),
         moisture_ave = as.numeric(MoistureAve),
         moisture_split = as.factor(Split),
         area = "NN") %>% 
  # replace 0 and -999 values with NA
  replace_with_na(replace = list(moisture_1 = 0, moisture_2 = 0,
                                 moisture_3 = 0, moisture_ave = 0)) %>% 
  replace_with_na(replace = list(moisture_1 = -999, moisture_2 = -999,
                                 moisture_3 = -999)) %>% 
  # gather 3 moisture samples to calculate average
  gather("sample", "value", c("moisture_1":"moisture_3")) %>% 
  arrange(home_winter_id, moisture_date) %>% 
  group_by(home_winter_id, moisture_date) %>% 
  mutate(moisture_ave = mean(value, na.rm = TRUE)) %>% 
  ungroup() %>% 
  # spread 3 moisture samples back to wide format
  spread(sample, value) %>% 
  select(home_winter_id, area, moisture_date, 
         moisture_1, moisture_2, moisture_3, 
         moisture_ave, moisture_split) %>% 
  replace_with_na(replace = list(moisture_ave = "NaN")) %>%
  filter(!is.na(moisture_ave)) %>% 
  arrange(home_winter_id, moisture_date)

# Combine locations and save
elders_moisture <- rbind(npt_moisture_initial, nn_moisture_initial) %>% 
  filter(!is.na(home_winter_id)) %>% 
  # lower range on instrument = 5%
  # therefore, changing values < 5 to 5
  mutate(moisture_1 = if_else(moisture_1 < 5, 5, moisture_1),
         moisture_2 = if_else(moisture_2 < 5, 5, moisture_2),
         moisture_3 = if_else(moisture_3 < 5, 5, moisture_3),
         moisture_ave = if_else(moisture_ave < 5, 5, moisture_ave),
         area = as.factor(area)) %>% 
  filter_at(vars(moisture_1:moisture_3), any_vars(!is.na(.))) %>% 
  arrange(area, home_winter_id, moisture_date) %>% 
  distinct(area, home_winter_id, moisture_date, .keep_all = T)

summary(elders_moisture)
```

# Join with demographic data to match on closest visit data
```{r}
elders_moisture_clean <- elders_moisture %>% 
  full_join(elders_demographics, by = c("area", "home_winter_id")) %>% 
  mutate(moisture_sample_date_diff = abs(moisture_date - sampling_date)) %>% 
  group_by(area, home_winter_id, sampling_visit) %>% 
  mutate(date_diff_rank = order(moisture_sample_date_diff)) %>% 
  ungroup() %>% 
  filter(date_diff_rank == 1) %>% 
  select(area, home_winter_id, sampling_visit, moisture_date,
         moisture_sample_date_diff, moisture_1:moisture_split)

# write_rds(elders_moisture_clean, paste0(file_path, "Output/elders_moisture_clean.rds"))
```
